# Non-Communicable Disease Burden (NCD)

```{r}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the non-communicable diseases data
data_ncd <- read_csv(here("data/burden-of-disease-ncd.csv"))
```

```{r}
# Filter the dataset to include only the three countries chosen by your group.
data_ncd_mini <- data_ncd %>% 
  filter(Entity%in%c("Ghana","Chile","Switzerland"))
#head(data_ncd_mini)
```

## Table of Estimates for NCD Burden Over Time

```{r}
# Here render a table for the DALY burden over time for the three countries. 
# You should pivot the data wider to show each country in a separate column.
data_ncd_wide <- data_ncd_mini %>%
  select(Year, Entity, `DALYs (Disability-Adjusted Life Years) - Non-communicable diseases - Sex: Both - Age: Age-standardized (Rate)`) %>% 
  pivot_wider(names_from = Entity,
              values_from = `DALYs (Disability-Adjusted Life Years) - Non-communicable diseases - Sex: Both - Age: Age-standardized (Rate)`)


# Use kable() from the knitr package to render the table.
  kable(data_ncd_wide, caption = "DALY Burden Over Time for Selected Countries")
```

## Plot Showing Trends in NCD Burden Over Time

```{r}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.
# NOTE: Use your long 'mini' data frame for plotting, not the wide data used for the table. 
ggplot(data=data_ncd_mini,
       aes(x=Year,
           y=`DALYs (Disability-Adjusted Life Years) - Non-communicable diseases - Sex: Both - Age: Age-standardized (Rate)`,
           color=Entity))+
  geom_line(linewidth=1)+
  labs(title = "Non-communicable diseases in children over time",
       color="Country",
       y="DALY Burden")+
  theme_minimal()

```

## Summary of NCD Burden Findings

Ghana has the overall highest burden on Communicable diseases in children compared to the other countries. Also, from the year 1990 to 2020, Switzerland and Chile have observed a significant decrease in the DALY burden among children, underscoring great improvement in the area while not so much improvement have been observed among children in Ghana.
